---
title: "R Notebook"
output: html_notebook
---

Read in libraries for work on this notebook
```{r}
library(dplyr)
library(ggplot2)
```


Read in the 2010 epi dataset and then adjust headings where first row entry is var names
```{r}
epi <- read.csv("/Users/paullee/Downloads/2010EPI_data.csv", header=TRUE)

View(epi)
names(epi) <- as.matrix(epi[1, ])
epi <- epi[-1, ]
epi[] <- lapply(epi, function(x) type.convert(as.character(x)))
View(epi)

summary(epi)
dim(epi)
```

Central Tendancies of Variables
```{r}
mode <- function(x) {
   uniq <- unique(x)
   uniq[which.max(tabulate(match(x, uniq)))]
}

#EPI Variable analysis
summary(epi$EPI)
fivenum(epi$EPI)
print("This is the mean of EPI")
mean(epi$EPI)
print("This is the median of EPI")
median(epi$EPI)
print("This is the mode of EPI")
mode(epi$EPI)

ggplot(epi, aes(x=EPI)) + 
  geom_histogram(binwidth = 2,colour = "blue") +
  ggtitle("Distribution of EPI")

qqnorm(epi$EPI, main="QQ Plot of EPI")
qqline(epi$EPI)

ecdf(epi$EPI)

#DALY Variable analysis
summary(epi$DALY)
fivenum(epi$DALY)
print("This is the mean of DALY")
mean(epi$DALY)
print("This is the median of DALY")
median(epi$DALY)
print("This is the mode of DALY")
mode(epi$DALY)

ggplot(epi, mapping = aes(x = DALY)) + 
  geom_histogram(binwidth = 2,colour = "blue") + 
  ggtitle("Distribution of DALY")

qqnorm(epi$DALY, main="QQ Plot of DALY")
qqline(epi$DALY)

#Observe relationship with each other
ggplot(epi, aes(x=EPI, y=DALY)) +
 geom_point(aes(color = EPI)) +
 geom_smooth(method ="lm") +
 coord_cartesian() +
 scale_color_gradient() +
 theme_bw() +
  ggtitle("DALY and EPI relationship")  

#Boxplot comparisons
boxplot(epi$EPI, epi$DALY, notch=TRUE, main="Boxplot distribution of EPI and DALY", xlab=("EPI and DALY"))

#Distribution comparison
qqplot(epi$EPI,epi$DALY, main="Daly and EPI QQ Plot")

# Correlation between epi and daly
x <- epi$EPI
y <- epi$DALY
#Correlation Coefficeint
cor(x, y)
```
```{r}
random_points <- sample_n(epi,5)
rand_points.df <- random_points %>%
  select(code, EPI, DALY)
rand_points.df

random_frac <- sample_frac(epi, .1)
rand_frac.df <- random_frac %>%
  select(code, EPI, DALY)
rand_frac.df

epi <- epi %>%
  arrange(desc(EPI)) %>%
  mutate(new_decs_EPI = EPI)

epi <- epi %>%
  arrange(desc(DALY)) %>%
  mutate(new_decs_DALY = DALY)

epi <- epi %>%
  mutate(double_EPI = EPI*2)

epi <- epi %>%
  mutate(double_DALY = DALY*2)

epi %>%
  summarise(avg_EPI = mean(EPI, na.rm=TRUE))

epi %>%
  summarise(avg_DALY = mean(DALY, na.rm = TRUE))
```

Visualize Envhealth and Ecosystem
```{r}
#Boxplot comparisons
boxplot(epi$ENVHEALTH, epi$ECOSYSTEM, notch=TRUE, main="Boxplot distribution of Envhealth and Ecosystem", xlab=("EnvHealth and Ecosystem"))

#Distribution comparison
qqplot(epi$ENVHEALTH, epi$ECOSYSTEM, main="Envhealth and Ecosystem QQ Plot")
```

Run multiple regression on all continous variables to observe relationship between DV and IV
```{r}
epi_lm.df <- epi[,6:164]
epi_lm <- lm(EPI ~ DALY + ENVHEALTH + ECOSYSTEM, data=epi_lm)
coefficients(epi_lm)
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

