---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
#install.packages("xgboost")
library(xgboost)
library(caret)
library(e1071) 
library(randomForest)
library(forecast)
library(Metrics)
library(Rtsne)
library(factoextra)
library(reshape2)
library(purrr)
library(dendextend)
library(cluster)
library(fastDummies)
library(dummies)
library(corrplot)
library(Boruta)
library(cowplot)
library(class)
library(ISLR)
library(RColorBrewer)
library(MLmetrics)
library(Metrics)
```

Read in Data frames

```{r}

wine_red <- read_csv2("/Users/kingg/Desktop/DataAnalytics/A7_Work/winequality-red.csv")
  
wine_white <- read_csv2("/Users/kingg/Desktop/DataAnalytics/A7_Work/winequality-white.csv")
```

Wine Exploration 
Modeling
  Classify red or white
  Classify quality (0-10)
  Regression Alcohol Content

```{r}
wine_red$type <- "red"
wine_white$type <- "white"

wine <- rbind(wine_red, wine_white)
colnames(wine) <- str_replace_all(colnames(wine), c(" "), "_")
wine[is.na(wine)] <- 0

ggplot(data=wine, aes(x=type, fill=as.factor(quality))) +
  ggtitle("Distribution of Red and White Wines") +
  geom_bar()

ggplot(data=wine, aes(x=type,y=alcohol, fill=as.factor(quality))) +
  ggtitle("Wine type by Alcohol Content") +
  geom_col()

#There are heavy outliers in alcohol content
wine <- wine %>%
  filter(alcohol <= fivenum(alcohol)[4])

ggplot(data=wine, aes(x=alcohol)) +
  geom_histogram(bins=5)
ggplot(data=wine, aes(reorder(type, alcohol), alcohol, fill=as.factor(quality))) +
  ggtitle("Boxplot of Alcohol Content by Wine Type") +
  coord_flip() +
  geom_boxplot()

ggplot(data=wine, aes(sample=alcohol)) +
  ggtitle("QQnorm of Alcohol Content") +
  geom_qq()

#Investigate the quality score of wine by each type
sort(unique(wine$quality)) #Supposed to be from 0-10 but 3-9
wine$quality <- as.factor(wine$quality)
ggplot(data=wine, aes(x=quality, fill=type)) +
  ggtitle("Distribution of wines by quality score") +
  geom_bar()

corr <- wine %>%
  select(-type)
corr <- sapply(corr, as.numeric)
res1 <- cor.mtest(corr, conf.level = .95)
corrplot(cor(corr), p.mat = res1$p, method = "color",
         sig.level = c(.001, .01, .05), pch.cex = .9,
         insig = "label_sig", pch.col = "white", order = "AOE", tl.cex=.5,  col = brewer.pal(n = 8, name = "RdYlBu"))

#shuffle the dataframe randomly before splitting to train and split
set.seed(12)
rows <- sample(nrow(wine))
wine <- wine[rows,]  

wine$type[which(wine$type == "red")] = 1
wine$type[which(wine$type == "white")] = 0
wine$type <- as.factor(wine$type)

train_index <- sample(1:nrow(wine), 0.8*nrow(wine))
test_index <- setdiff(1:nrow(wine), train_index)

train <- wine[train_index,]
test <- wine[test_index,]

```

Clustering using Gower Distance and K-Mediod then visualizing using rtsne dimension reduction

```{r}
#Gower distance to clustering to see if any underlying groups exists in data
cluster_df <- wine %>%
  mutate_if(is.character, funs(as.factor(.)))
distance <- daisy(cluster_df, metric="gower")
gower_matrix <- as.matrix(distance)

cluster_df[which(gower_matrix == min(gower_matrix[gower_matrix != min(gower_matrix)]), arr.ind = TRUE)[1, ], ]

cluster_df[which(gower_matrix == max(gower_matrix[gower_matrix != max(gower_matrix)]), arr.ind = TRUE)[1, ], ]

sil_width <- c(NA)
for(i in 2:8){  
  pam_fit <- pam(distance, diss = TRUE, k = i)  
  sil_width[i] <- pam_fit$silinfo$avg.width  
}
plot(1:8, sil_width,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:8, sil_width)
#Select 4 clusters maximum sil coefficient and not big increase to 5 clusters
k <- 4
#k-mediods model for clustering based on optimal num clusters for max sil coefficient
pam_fit <- pam(distance, diss = TRUE, k)

#Summary profiles based on grouping rows to their appropriate clustered group
pam_results <- cluster_df %>%
  mutate(cluster = pam_fit$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))
pam_results$the_summary

#Rtsne dimension redution visualization
tsne_obj <- Rtsne(distance, is_distance = TRUE)
tsne_data <- tsne_obj$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering))

ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster)) +
  ggtitle("Rtsne cluster visualization")

ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = wine$quality)) +
  ggtitle("Rtsne cluster visualization")
```

KNN Clustering 

```{r}
knn_train <- train %>%
  select(-type)
knn_test <- test %>%
  select(-type)



knn_train$quality <- as.factor(knn_train$quality)
knn_test$quality <- as.factor(knn_test$quality)

knn_train_x <- knn_train[,-12]
knn_train_x <- sapply(knn_train_x, as.numeric)
knn_train_y <- knn_train$quality

knn_test_x <- knn_test[,-12]
knn_test_x <- sapply(knn_test_x, as.numeric)
knn_test_y <- knn_test$quality


error=c()
for (k in 1:200){
  predicted = knn(train = scale(knn_train_x),
                  test = scale(knn_test_x),
                  cl = knn_train_y,
                  k=k)
  error[k]=mean(knn_test_y != predicted)
}

plot(error, type="b",ylab="Error",xlab="Number of Neighbors", main="Error over  K")
best6 <- head(sort(error, decreasing=FALSE))
error
print(max(which(error == best6), error))
#k=59

final_pred <- knn(train = scale(knn_train_x),
                  test = scale(knn_test_x),
                  cl = knn_train_y,
                  k=59)
confusionMatrix(final_pred, knn_test_y)
```



Simple Binary Classification problem look at predicting if something is white or wine

RandomForest

```{r}
rf <- randomForest(type~., data = train, importance = TRUE)
summary(rf)

rf_pred <- predict(rf, test, type="class")

confusionMatrix(rf_pred, test$type)
varImpPlot(rf)
```


Multiclass using XGBoost to predict quality of wine more complex problem 

```{r}
#XGBoost requires that our y labels or classes start from 0 and increase sequentially in our case quality should go from 0-10 but we only have values from 3 4 5 6 7 8 9.
xgb_wine <- wine %>%
  filter(quality %in% c(5,6,7))
xgb_wine$quality <- as.numeric(levels(xgb_wine$quality))[xgb_wine$quality]
unique(xgb_wine$quality)

#quality <- c(3,4,5,6,7,8,9)
quality <- c(5,6,7)
class_encoder <- 0
for (i in quality){
  for (j in 1:nrow(xgb_wine)){
    if(xgb_wine$quality[j] == i){
      xgb_wine$quality[j] = class_encoder
    }
  }
  class_encoder <- class_encoder +1
}

#all categoricals must become numeric for xgboost
xgb_matrix <- data.matrix(xgb_wine)

#Xgboost for multiclass using sgboost library requires matrix input
xgb_quality <- xgb_matrix[,"quality"]

xgb_vars <- xgb_matrix[,-12]

#Create xgb class object for modeling 
xgb_data <- xgb.DMatrix(data = xgb_matrix, label = xgb_quality)

train_index <- sample(1:nrow(xgb_wine), 0.8*nrow(xgb_wine))
test_index <- setdiff(1:nrow(xgb_wine), train_index)

#Create train matrix data with associated labels
xgb_train <- xgb_vars[train_index,]
xgb_train_quality <- xgb_quality[train_index]
xgb_train_matrix <- xgb.DMatrix(data=xgb_train, label=xgb_train_quality)

#Create test matrix data with associated labels
xgb_test <- xgb_vars[test_index,]
xgb_test_quality <- xgb_quality[test_index]
xgb_test_matrix <- xgb.DMatrix(data=xgb_test, label=xgb_test_quality)

num_quality <- length(unique(xgb_wine$quality))
xgb_params <- list("objective" = "multi:softprob","eval_metric" = "mlogloss","num_class" = num_quality)
nround <- 500 
cv.nfold  <- 50

cv_model <- xgb.cv(params = xgb_params,
                   data = xgb_train_matrix, 
                   nrounds = nround,
                   nfold = cv.nfold,
                   verbose = FALSE,
                   prediction = TRUE)

xgb_pred<- data.frame(cv_model$pred) %>%
  mutate(max_prob = max.col(., ties.method = "last"),label = xgb_train_quality + 1)
head(xgb_pred)

confusionMatrix(factor(xgb_pred$max_prob), factor(xgb_pred$label), mode = "everything")

ggplot(data=xgb_pred, aes(x=as.factor(max_prob))) +
  geom_bar()

wine_quality_reduce <- wine %>%
  filter(quality %in% c(5,6,7))
ggplot(data=wine_quality_reduce, aes(x=quality)) +
  geom_bar()
```

  
Banking
Modeling
  Classify customers who have accepted term deposit or not (1-0)
  
```{r}
absent <- read_csv2("/Users/kingg/Desktop/DataAnalytics/A7_Work/Absenteeism_at_work.csv")
head(absent)

colnames(absent) <- str_replace_all(colnames(absent), c(" "), "_")
colnames(absent)[21] <- "absent_hours"
colnames(absent)

ggplot(data=absent, aes(x=absent_hours, fill=Reason_for_absence)) +
  geom_histogram(bins=10) +
  ggtitle("Distribution of hours absent by Reason for absences")

unique(absent$Reason_for_absence)
absent$Reason_for_absence <- as.factor(absent$Reason_for_absence)

ggplot(data=absent, aes(reorder(Reason_for_absence,absent_hours), absent_hours, fill=as.factor(Disciplinary_failure))) +
  geom_boxplot(color="blue") +
  coord_flip() +
  ggtitle("Boxplots of Hours absent by Reason codes and Disciplinary_failures")

ggplot(data=absent, aes(reorder(Reason_for_absence, -absent_hours, sum), absent_hours, fill=as.factor(Education))) +
  geom_col() +
  ggtitle("Distribution of Hours absent by Reason codes and Education")

unique(absent$Seasons)
unique(absent$Month_of_absence) #How can there be month 0? also these values with month 0 have a absent hour value of 0 so consider removing
unique(absent$Day_of_the_week) #sat=7 and mon =1 so values from (2-6) repersent working week day

absent <- absent %>%
  filter(Month_of_absence > 0)
absent$Month_abb <- month.abb[absent$Month_of_absence]

absent_ts <- absent %>%
  mutate(month_day = paste(Month_abb,Day_of_the_week, sep="-")) %>%
  select(month_day, absent_hours, Seasons, Reason_for_absence)

ggplot(data=absent_ts, aes(x=month_day, y=absent_hours, fill=absent$Reason_for_absence)) +
  geom_col() +
  theme(axis.text.x = element_text(face="bold",size=5, angle=90)) +
  ggtitle("Absent Hours by Time and Reason for Absence codes")

sea1 <- absent_ts %>%
  filter(Seasons == 1)
sea2 <- absent_ts %>%
  filter(Seasons == 2)
sea3 <- absent_ts %>%
  filter(Seasons == 3)
sea4 <- absent_ts %>%
  filter(Seasons == 4)

s1 <- ggplot(sea1, aes(x=month_day, y=absent_hours, fill=Reason_for_absence)) +
  geom_col() +
  theme_half_open(12) +
  theme(plot.margin= ggplot2::margin(6,0,6,0)) +
  theme(axis.text.x = element_text(face="bold",size=5, angle=90)) +
  ylab(NULL)
s2 <- ggplot(sea2, aes(x=month_day, y=absent_hours, fill=Reason_for_absence)) +
  geom_col() +
  theme_half_open(12) +
  theme(plot.margin= ggplot2::margin(6,0,6,0)) +
  theme(axis.text.x = element_text(face="bold",size=5, angle=90)) +
  ylab(NULL)
s3 <- ggplot(sea3, aes(x=month_day, y=absent_hours, fill=Reason_for_absence)) +
  geom_col() +
  theme_half_open(12) +
  theme(plot.margin= ggplot2::margin(6,0,6,0)) +
  theme(axis.text.x = element_text(face="bold",size=5, angle=90)) +
  ylab(NULL)
s4 <- ggplot(sea4, aes(x=month_day, y=absent_hours, fill=Reason_for_absence)) +
  geom_col() +
  theme_half_open(12) +
  theme(plot.margin= ggplot2::margin(6,0,6,0)) +
  theme(axis.text.x = element_text(face="bold",size=5, angle=90)) +
  ylab(NULL)

legend <- get_legend(s1 + 
    guides(shape = guide_legend(override.aes = list(size = 5))) +
    guides(color = guide_legend(override.aes = list(size = 5))) +
    theme(legend.title = element_text(size = 7),legend.text = element_text(size = 5)) +
      theme(legend.box.margin = ggplot2::margin(0,0,0,12)))

group_plot <- plot_grid(s1 + theme(legend.position="none"),s2 + theme(legend.position="none"),s3 + theme(legend.position="none"),s4 + theme(legend.position="none"), labels=c("Season1","Season2","Season3","Season4"),label_size=7,hjust=-1)

plot_grid(group_plot, legend, ncol = 2, rel_widths = c(4,1), align=c("hv"))

ggplot(data=absent) +
  geom_qq(aes(sample=absent_hours)) +
  ggtitle("QQnorm of Absent hours")

absent <- absent %>%
  select(-Month_abb)

boruta <- absent
boruta <- as.data.frame(boruta)

dummy_cols <- c("Reason_for_absence","Month_of_absence","Day_of_the_week","Seasons","Disciplinary_failure","Education","Son","Social_drinker","Social_smoker","Pet")

boruta[dummy_cols] <- lapply(boruta[dummy_cols], factor)

set.seed(101)
boruta_train <- Boruta(absent_hours~. -ID, data=boruta, doTrace=2)
plot(boruta_train)
boruta_final <- TentativeRoughFix(boruta_train)
plot(boruta_final)
boruta_vars <- attStats(boruta_final)

index_holders <- which(boruta_vars$decision == "Confirmed", arr.ind=T)
model_vars <- row.names(boruta_vars[index_holders,])

attStats(boruta_final)

model_vars[6] <- "Work_load_Average/day"

absent <- absent %>%
  select(model_vars, absent_hours)

new_dummy_cols <- c("Reason_for_absence","Seasons","Disciplinary_failure","Education")
colnames(absent)

#Creating train/test and dummies
absent_dum <- data.frame(dummy_cols(absent,select_columns=new_dummy_cols,remove_first_dummy = TRUE,remove_selected_columns = TRUE))

absent_train_index <- sample(1:nrow(absent_dum), 0.8*nrow(absent_dum))
absent_test_index <- setdiff(1:nrow(absent_dum), absent_train_index)

absent_train <- absent_dum[absent_train_index,-c(1)]
absent_test <- absent_dum[absent_test_index,-c(1)]


```

Linear Regression model for Prediciting hours absent
```{r}
absent_lm <- lm(absent_hours~., data = absent_train)
summary(absent_lm)

absent_lm_pred <- predict(absent_lm, absent_test)

round(absent_lm_pred)
absent_test$absent_hours

plot(round(absent_lm_pred),absent_test$absent_hours)
abline(0,1)

lm_mse <- mean((round(absent_lm_pred)-absent_test$absent_hours)^2)
MAPE(absent_lm_pred, absent_test$absent_hours, na.rm=TRUE)
smape(absent_lm_pred, absent_test$absent_hours)
forecast::accuracy(absent_test$absent_hours, absent_lm_pred)


absent_rf <- randomForest(absent_hours~., data = absent_train, importance = TRUE)
summary(absent_rf)

absent_rf_pred <- predict(absent_rf, absent_test)
plot(round(absent_rf_pred),absent_test$absent_hours)
abline(0,1)

absent_rf_mse <-  mean((round(absent_rf_pred)-absent_test$absent_hours)^2)
smape(absent_rf_pred, absent_test$absent_hours)
forecast::accuracy(absent_rf_pred, absent_test$absent_hours)
```




































